The Blessing of FedAvg's Linear Speedup in Federated Learning

### 2020-05-29 source/sec_exp.tex
experiments appendix details
add conclusion

\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, we
conduct experiments on both synthetic datasets and real-world dataset
w8a~\cite{platt1998fast} $(d=300, n=49749)$. We consider the distributed
objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
each local device $k$ includes three cases: 1) \textbf{Strongly convex
objective}: the regularized binary logistic regression problem, $F_k(\vw) =
\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \lambda
\|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
1e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
problem without regularization. $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k}
\log( 1+ \exp(-y_i^k \vw^T\vx_i^k)$. 3) \textbf{Overparameterized setting}:
the linear regression problem without adding noise to the label, $F_k(\vw) =
\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  In the first
two cases, we use the w8a dataset to conduct empirical evaluation. 
In the linear regression problem, we use features from w8a dataset and 
generate a ground truth $w$ to label each instance, which satisfies the
overparameterized setting.

% where we generated i.i.d. data $\vx_i^k$ for all devices. At each communication round, all devices participated
% in synchronization. 
\textbf{Linear speedup.} To verify the linear speedup results in Theorem~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}, we evaluate the number of iterations needed to reach
$\epsilon$-accuracy in three objectives. We initialize all runs with $w_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
the best seed results for each configuration $(E, K)$. In each configure, we
extensive search the stepsize from $\min(\eta_0, \frac{nc}{1 + t})$, where
$\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
take the values $c = 2^u \forall i \in \ZZ$. For the detailed experimental setup,
please refer to the appendix. As the results shown in Figure~\ref{fig:speedup},
the number of iterations to reach the target accuracy $\epsilon$ decreasing 
as the number of workers increase.

@@

@@

\begin{figure}
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
& 
\includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
	\end{tabular}
\caption{The linear speedup convergence of Accelerated FedAvg w.r.t the number of workers. }
\label{fig:nesterov}
\end{figure}

>>>>

\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, we
conduct experiments on both synthetic datasets and real-world dataset
w8a~\cite{platt1998fast} $(d=300, n=49749)$. We consider the distributed
objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
each local device $k$ includes three cases: 1) \textbf{Strongly convex
objective}: the regularized binary logistic regression problem, $F_k(\vw) =
\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \lambda
\|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
1e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
problem without regularization. $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k}
\log( 1+ \exp(-y_i^k \vw^T\vx_i^k)$. 3) \textbf{Overparameterized setting}:
the linear regression problem without adding noise to the label, $F_k(\vw) =
\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  In the first
two cases, we use the w8a dataset to conduct empirical evaluation. 
In the linear regression problem, we use features from w8a dataset and 
generate a ground truth $w$ to label each instance, which satisfies the
overparameterized setting. For the detailed experimental setup and
hyperparameter setting, please refer to the section~\ref{sec:expsupp} in the
appendix.

% ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
\textbf{Linear speedup.} To verify the linear speedup results in Theorems
\lkxcom{cite sgd theorems}, we evaluate the number of iterations needed to reach
$\epsilon$-accuracy in three objectives. We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
the best seed results for each configuration $(E, K)$. In each configure, we
extensive search the stepsize from $\min(\eta_0, \frac{nc}{1 + t})$, where
$\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
the number of iterations to reach the target accuracy $\epsilon$ decreasing 
as the number of workers increase.

@@

\textbf{Nesterov accelerated FedAvg.} In Figure~\ref{fig:nesterov}, 
report the convergence of Nesterov accelerated FedAvg. The empirical 
observation is aligned with our theorems \lkxcom{cite accelerated theorems} that the acclerated version of FedAvg can also achieve the lienar speedup
w.r.t the number of nodes.

@@

\begin{figure}
\centering
\begin{tabular}{ccc}
\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
& 
\includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
	\end{tabular}
\caption{The linear speedup convergence of Nesterov accelerated FedAvg w.r.t the number of workers. }
\label{fig:nesterov}
\end{figure}
